---
title: "hw6.qmd"
format: html
---

# Homework 6
## Task 1
1. What is the purpose of the lapply() function? What is the equivalent purrr function?
    - lapply applies a function to each element of a list
    - purrr equivalent function is `map ()`
2. Suppose we have a list called my_list. Each element of the list is a numeric data frame (all columns
are numeric). We want use lapply() to run the code cor(numeric_matrix, method = "kendall")
on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify
method = "kendall" when calling lapply())
```{r}
library(purrr)
help("matrix")
numeric_matrix <- list(matrix(rnorm(25), ncol=5), matrix(rnorm(25), ncol=5))
lapply_example <- lapply(numeric_matrix, FUN = cor,  method = "kendall")
```

3. What are two advantages of using purrr functions instead of the BaseR apply family?
    1. automatic handling of data types and type consistency through the map_* functions
    2. the same inconsistent approach to function arguments and implementations in BaseR. 
        Conway's Law of Anarchy is the best description. They built a system that matched
        the communication structure of the organization. 
4. What is a side-effect function?
    - A side effect function is a function that represents the underlying data e.g. `print`
5. Why can you name a variable sd in a function and not cause any issues with the sd function?
    - Namespace and scoping. Variables inside of functions are scoped to the local
      function. This means the variable will only exist during the function runtime. Once the function             returns, the `sd` variable will be deallocated from memory. Since the function `sd` is a different           object in a separately scoped namespace (`stats`) than the variable `sd`, R will only look for a more        global scope if no locally scoped `sd` object does not exist. 
      
      
## Task 2 - Writing R Functions

### 1. `getRMSE` (vector of responses,vector of predictions)
```{r}
getRMSE <- function (response, predicted){
  sqrt(mean((response - predicted)^2))
}
```
#### Handle NA as an argument to `getRMSE`
```{r}
getRMSE <- function (response, predicted, ...){
  sqrt(mean((response - predicted)^2, ...))
}
```
### 2. Test Function
#### Test all numeric values
```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
rmse_tst <- getRMSE(resp, pred)
print(rmse_tst)

```

#### Test NA values present
```{r}
rando_index_1 <- sample(length(resp),1)
rando_index_2 <- sample(length(pred),1)
resp[rando_index_1] <- NA_real_
pred[rando_index_2] <- NA_real_
rmse_tst <- getRMSE(resp, pred, na.rm = TRUE)
print(rmse_tst)
```
#### Test function calls (na.rm = FALSE, na.rm = TRUE)
```{r}
# assert na.rm = TRUE returns a number
rmse_tst <- getRMSE(resp, pred, na.rm = TRUE)
print(rmse_tst)

# assert na.rm = FALSE returns NA
rmse_tst <- getRMSE(resp, pred, na.rm = FALSE)
print(rmse_tst)
```


### 3. `getMAE`: function (<vector of responses>,<vector of predictions>)
```{r}
getMAE <- function (response, predicted, ...){
  mean(abs(response - predicted), ...)
}
```

### 4. Test MAE
#### Test all numeric values
```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
mae_tst <- getMAE(resp,pred)
print(mae_tst)

```
#### Test NA values present
```{r}
rando_index_1 <- sample(length(resp),1)
rando_index_2 <- sample(length(pred),1)
resp[rando_index_1] <- NA_real_
pred[rando_index_2] <- NA_real_
mae_tst <- getMAE(resp,pred)
print(mae_tst)
```

#### Test function calls (na.rm = FALSE, na.rm = TRUE)
```{r}
# assert na.rm = TRUE returns a number
mae_tst <- getMAE(resp, pred, na.rm = TRUE)
print(mae_tst)

# assert na.rm = FALSE returns NA
mae_tst <- getMAE(resp, pred, na.rm = FALSE)
print(mae_tst)
```

### 5. `snoop` : wrapper function (return either or both of `getRMSE` and `getMAE`)
```{r}
snoop <- function(resp,
                  pred,
                  metrics = c("rmse","RMSE","getRMSE","mae","MAE","getMAE"), ...){
  if (!(is.numeric(resp) & 
        is.vector(resp)  & 
        is.atomic(resp)  &
        is.numeric(pred) &
        is.atomic(pred)  &
        is.vector(pred))){
    stop("Prediction and Response should be numeric vectors")
  }
  # dotdotdot <- names(list(...))
  # print(dotdotdot)
  metrics <- tolower(metrics)
  result <- list(rmse = NULL, mae = NULL)
  
  if(is.null(metrics)){
    result$rmse <- getRMSE(resp, pred, ...)
    result$mae  <- getMAE(resp, pred, ...)
  }
  else {
    if("rmse" %in% metrics | "getrmse" %in% metrics){
      result$rmse <- getRMSE(resp, pred, ...)
    }
    if("mae" %in% metrics | "getmae" %in% metrics){
      result$mae <- getMAE(resp,pred, ...)
    }
    return(result)
  }
}
```

#### Test all numeric values
```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))

rmse_tst <- snoop(resp, pred, "rmse")
mae_tst <- snoop(resp, pred, "MAE")
both_tst <- snoop(resp, pred)
```

#### Test NA values present
```{r}
rando_index_1 <- sample(length(resp),1)
rando_index_2 <- sample(length(pred),1)
resp[rando_index_1] <- NA_real_
pred[rando_index_2] <- NA_real_

rmse_na_tst <- snoop(resp, pred, "rmse", na.rm = TRUE)
mae_na_tst <- snoop(resp, pred, "MAE", na.rm = TRUE)
both_na_tst <- snoop(resp, pred,na.rm = TRUE)
```

#### Test function calls (default: rmse = TRUE, mae = TRUE)
```{r}
# df <- data.frame(col1 = 0.1, col2 = 0.2)
# df_tst <- snoop(resp, df)
# df_tst <- snoop(df, pred, "rmse")
```


## Task 3 - Query an API and Tidy-Style Function

### 1. Return info about a topic
### 2. Parse returned info, grab articles element

```{r, echo=FALSE}
library(tidyverse)
library(httr)
library(jsonlite)
library(usethis)
```

```{r}
#store key
# usethis::edit_r_environ()
readRenviron("~/.Renviron")
news_api_key <- Sys.getenv("NEWS_API_KEY")

query <- URLencode("FEMA+Militia+Chimney Rock")
from_date <- "2024-10-01"
to_date <- "2024-10-15"
url <- paste0("https://newsapi.org/v2/everything?q=", query,
              "&from=", from_date,
              "&to=", to_date,
              "&sortBy=popularity&apiKey=", news_api_key)

# print(url)
news_info <-GET(url) |>
  content("text") |>
  fromJSON(flatten = TRUE, simplifyDataFrame = TRUE) |>
  pluck("articles")

news_info <- as.tibble(news_info)
```

### 3. Write function to call the API
```{r}
news_query <- function(subject, start_date, api_key){
  
}
```

